---
title: "Frequentist Approach to Green Sturgeon Rheotaxis Data"
author: "Myfanwy Johnston"
date: "August 3, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

These experiments were originally conceived as multiple Latin Squares, with fish randomly assigned to a square (six squares, four treatments per square, one fish per treatment, for a total of 24 fish).  However, one fish had to be re-used, and so had "membership" in two of the six Latin Squares, changing the total to 23 fish.

Another concern was that the four treatments (Above, Below, Light, and Dark) were run very close together, with at most an hour between two treatments.  This raised questions about whether we might observe carryover effects from the previous treatment. 

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

# Why we went with Bayesian Approach instead:

Because each fish underwent each treatment only once in experiment 1 (with the exception of the one fish that underwent each treatment twice), we were unable to include a fish*treatment random effect in the set of candidate models.  However, from what we know about green sturgeon data, the interaction between individuals and treatment was an important factor to include in the models.  The Bayesian approach allowed us to incorporate this important variable as part of the varying slopes and intercepts model.


